Conversion of FFCS ICE fleet to EV: a data-driven approach

### 2020-11-12 Chapter7/7_1_intro.tex
chapt

I start by leverage the data collected from more than 2 months of rentals in 2017 with the software described in chapter \ref{chap:2_dataset}. I remind to the characterization done in sections \ref{sec:3_2_results} and \ref{sec:5_3_mh_placement}

@@

Reminding that section \ref{sec:6_5_freefloating} shows that placing the charging stations in those areas where cars stay parked for long periods performs worse than a totally agnostic random placement. Instead, placing charging station in those areas where cars are frequently parked even for short periods guarantees  better performance.

@@

After discussing related work in Section~\ref{sec:7_2_related}, I present the methodology I adopted to collect data and characterise the data-set in Section~\ref{sec:7_3_data}.
 In  Section~\ref{sec:7_4_modelling} I describe the simulation model, its parameters and metrics of interest. Section~\ref{sec:7_5_placement} presents the different placement heuristics and the algorithms I design to optimise the placement.  Section~\ref{sec:7_6_herustic_placement} discusses the impact of simple charging stations placement policies and return policies, while Section~\ref{sec:7_7a_heuristic_vs_opt} reports the results of the optimisation and their validation. Section~\ref{sec:7_8_discussion} discusses limitations and future work, before drawing conclusions in Section~\ref{sec:7_9_conclusion}.

>>>>

I start by leverage the data collected from more than 2 months of rentals in 2017 with the software described in chapter \ref{chap:2_dataset}. I remind to the characterization done in sections \ref{sec:3_4_results} on Turin and \ref{sec:5_3_mh_placement} studying in detail Turin, Vancouver and Berlin.

@@

Reminding that section \ref{sec:6_6_freefloating} shows that placing the charging stations in those areas where cars stay parked for long periods performs worse than a totally agnostic random placement. Instead, placing charging station in those areas where cars are frequently parked even for short periods guarantees  better performance.

@@

After discussing related work in Section~\ref{sec:7_2_related}, I present the methodology I adopted to collect data and characterise the data-set in Section~\ref{sec:7_3_data}.
 In  Section~\ref{sec:7_4_modelling} I describe the simulation model, its parameters and metrics of interest. Section~\ref{sec:7_5_placement} presents the different placement heuristics and the algorithms I design to optimise the placement.  Section~\ref{sec:7_7a_opt} discusses the impact of simple charging stations placement policies and return policies, while Section~\ref{sec:7_7b_needed} reports the results of the optimisation and their validation. Section~\ref{sec:7_8_discussion} discusses limitations and future work, before drawing conclusions in Section~\ref{sec:7_9_conclusion}.

### 2020-11-12 Chapter7/7_7a_heuristic_vs_optimized.tex
chapt

In section \mc{indicare sezione},  I pointed out how \textit{Num Parking} placement heuristic works better than the other two.
Weighting also charges and re-routing, the \textit{Hybrid} policy shows better performances than \textit{Needed}. For this reason, in this section, I focus on the \textit{Hybrid} return policy with charging stations covering less than 15\% of the zones. I further optimise this scenario by running the meta-heuristic placement algorithms, and comparing the results with the \textit{Num Parking} placement. 
Optimisation with \textit{Needed} return policies are briefly discussed in \ref{sec:7_7b_needed}, where very similar results are obtained.

@@

\begin{figure}[th]
    \centering     %%% not \center
    \subfloat[][Percentage of infeasible trips.  Y-Axis is logarithmic.]
    {
      \includegraphics[width=0.45\textwidth]{figures/Hybrid_Deaths.pdf}
        \label{fig:7_7a_optimized_deaths}
    }
    \subfloat[][Average walked distance.]
    {
        \includegraphics[width=0.45\textwidth]{figures/Hybrid_TravelWithPenlaty.pdf}
        \label{fig:7_7a_optimized_penaly}
    }    
    \label{fig:7_7a_optimized_metrics}
    \caption{Objective metrics to minimise in the optimisation - with \textit{Hybrid} return policy.}
\end{figure}

\begin{figure*}[t!]
	\begin{center}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Hybrid_AmountRechargePerc.pdf}
			\caption{Average state of charge.}
			\label{fig:7_7a_recharge}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/AvgSOC_comparison}
			\caption{Average state of charge.}
			\label{fig:7_7a_asoc_Needed}
		\end{subfigure}

@@

Figure.~\ref{fig:7_7a_optimized_metrics} reports the two target metrics, for all the optimised configurations.
Firstly, in figure.~\ref{fig:7_7a_optimized_deaths} I compare the infeasible trip percentage. \textit{Num Parking} solution (blue line) has already good performance, reaching 0 with 4.2\% of the equipped zones. 
The \textit{Local Search} (orange line) and the \textit{Genetic} (green line) algorithms are able to further reduce the minimum percentage of zone to equip to guarantee no infeasible trips: 3.8\% by the \textit{Local Search}, and to 3.5\% with the \textit{Genetic} algorithm, $Z=10$ and $Z=9$ zones respectively.

Figure.~\ref{fig:7_7a_optimized_penaly} reports the walked distance. Focusing in the feasible region, the \textit{Genetic} algorithm confirms the best performance, reducing the distance from more than 200\,m to 136\,m when 4.2\% of the zones are equipped with charging stations, and reaching just 30\,m at 13\%.

In figure.~\ref{fig:7_7a_hybrid_optimization} I further study the new solutions on other metrics.
Figure \ref{fig:7_7a_recharge} reports the percentage of trips ending in a charging station. The more charges are performed, the more time the customer has to spend time plugging/unplugging the car. By minimising the walked distance, I also reduce this metric, since a trip ending with a charge corresponds to 150\,m of penalty.
Here, the \textit{Local Search} follows the same trend as the \textit{Num parking} with a strong rise. The \textit{Genetic} algorithm shows much better results, from 9\% to 14\% of trips ending with a charge -- half of those found with other solutions. 
This improvement highlights the better placement of the charging stations.
Focus now on figure. \ref{fig:7_7a_asoc}, which reports the average state of charge of the car battery. 
No major difference are observed here, with all curves almost overlapping up to 7\% of the zones.

@@

To give a feeling about the differences in the solutions found by different algorithms, figure.~\ref{fig:7_7a_maps} reports the solutions obtained with 7\% of the zones equipped with charging stations (i.e., $Z=18$ zones).

\textit{Num parking} solution, figure~\ref{fig:7_7a_hm_max-parking_3}, places most of the charging stations in downtown area and near the main train stations. 
\textit{Local Search}, figure~\ref{fig:7_7a_hm_local3}, still has many zones in common with \textit{Num parking}, the solution it started from. It just spreads some charging station to cover also some remote zones.
The \textit{Genetic} algorithm, figure.~\ref{fig:7_7a_hm_genetico3}, shows very few zones in common with \textit{Num parking}. Charging stations are spread all over the city, still with more density in the city centre.

>>>>

In section \ref{sec:6_6_freefloating},  I pointed out how \textit{Num Parking} placement heuristic works better than the other two.
Weighting also charges and re-routing, the \textit{Hybrid} policy shows better performances than \textit{Needed}. For this reason, in this section, I focus on the \textit{Hybrid} return policy with charging stations covering less than 15\% of the zones. I further optimise this scenario by running the meta-heuristic placement algorithms, and comparing the results with the \textit{Num Parking} placement. 
Optimisation with \textit{Needed} return policies are briefly discussed in \ref{sec:7_7b_needed}, where very similar results are obtained.

@@

\begin{figure}[th]
    \centering     %%% not \center
    \subfloat[][Percentage of infeasible trips.  Y-Axis is logarithmic.]
    {
      \includegraphics[width=0.45\textwidth]{figures/Hybrid_Deaths.pdf}
        \label{fig:7_7a_optimized_deaths}
    }
    \subfloat[][Average walked distance.]
    {
        \includegraphics[width=0.45\textwidth]{figures/Hybrid_TravelWithPenlaty.pdf}
        \label{fig:7_7a_optimized_penaly}
    }    
    \label{fig:7_7a_opt_metrics}
    \caption{Objective metrics to minimise - with \textit{Hybrid} return policy.}
\end{figure}

\begin{figure*}[t!]
	\begin{center}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Hybrid_AmountRechargePerc.pdf}
			\caption{Average state of charge.}
			\label{fig:7_7a_recharge}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/AvgSOC_comparison}
			\caption{Average state of charge.}
			\label{fig:7_7a_asoc}
		\end{subfigure}

@@

Figure~\ref{fig:7_7a_opt_metrics} reports the two target metrics, for all the optimised configurations.
Firstly, in figure~\ref{fig:7_7a_optimized_deaths} I compare the infeasible trip percentage. \textit{Num Parking} solution (blue line) has already good performance, reaching 0 with 4.2\% of the equipped zones. 
The \textit{Local Search} (orange line) and the \textit{Genetic} (green line) algorithms are able to further reduce the minimum percentage of zone to equip to guarantee no infeasible trips: 3.8\% by the \textit{Local Search}, and to 3.5\% with the \textit{Genetic} algorithm, $Z=10$ and $Z=9$ zones respectively.

Figure~\ref{fig:7_7a_optimized_penaly} reports the walked distance. Focusing in the feasible region, the \textit{Genetic} algorithm confirms the best performance, reducing the distance from more than 200\,m to 136\,m when 4.2\% of the zones are equipped with charging stations, and reaching just 30\,m at 13\%.

In figure~\ref{fig:7_7a_hybrid_optimization} I further study the new solutions on other metrics.
Figure \ref{fig:7_7a_recharge} reports the percentage of trips ending in a charging station. The more charges are performed, the more time the customer has to spend time plugging/unplugging the car. By minimising the walked distance, I also reduce this metric, since a trip ending with a charge corresponds to 150\,m of penalty.
Here, the \textit{Local Search} follows the same trend as the \textit{Num parking} with a strong rise. The \textit{Genetic} algorithm shows much better results, from 9\% to 14\% of trips ending with a charge -- half of those found with other solutions. 
This improvement highlights the better placement of the charging stations.
Focus now on figure \ref{fig:7_7a_asoc}, which reports the average state of charge of the car battery. 
No major difference are observed here, with all curves almost overlapping up to 7\% of the zones.

@@

To give a feeling about the differences in the solutions found by different algorithms, figure~\ref{fig:7_7a_maps} reports the solutions obtained with 7\% of the zones equipped with charging stations (i.e., $Z=18$ zones).

\textit{Num parking} solution, figure~\ref{fig:7_7a_hm_max-parking_3}, places most of the charging stations in downtown area and near the main train stations. 
\textit{Local Search}, figure~\ref{fig:7_7a_hm_local3}, still has many zones in common with \textit{Num parking}, the solution it started from. It just spreads some charging station to cover also some remote zones.
The \textit{Genetic} algorithm, figure~\ref{fig:7_7a_hm_genetico3}, shows very few zones in common with \textit{Num parking}. Charging stations are spread all over the city, still with more density in the city centre.

### 2020-11-12 Chapter7/7_7b_needed.tex
chapt

 \begin{figure}[h]
     \centering     %%% not \center
    \subfloat[][Percentage of infeasible trips. Y-Axis is logarithmic.]
    {
        \includegraphics[width=0.45\textwidth]{figures/Needed_Deaths.pdf}
        \label{fig:7_7a_optimized_deaths_Needed}
    }     
     \subfloat[][Walked distance, averaged over all trips.]
     {
         \includegraphics[width=0.45\textwidth]{figures/Needed_TravelWithPenlaty.pdf}
         \label{fig:7_7a_wwd_Needed}
     }
     \caption{Objective metrics to minimise in the optimisation - with \textit{Needed} return policy}
    \label{fig:7_7a_optimized_metrics_needed}
 \end{figure}

Here I briefly report the results for the optimisation experiments of the \textit{Needed} policy.
I followed the same procedure explained in Sec.~\ref{sec:7_7a_opt} for the \textit{Hybrid} return policy. As in that case, the genetic algorithm is able to largely optimise the solution, as reported in figure~\ref{fig:7_7a_optimized_metrics_needed}, with \textit{local searches} stuck in local minima. 
In particular, for the walked distance (figure~\ref{fig:7_7a_wwd_Needed}), the genetic algorithm is able lower it from 136\,m to 45\,m at 13\% of zones. Still, it doesn't reach the performance of the \textit{Hybrid} policy, i.e., 30\,m at 13\% of zones.

\begin{figure*}[t!]
	\begin{center}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Needed_AmountRechargePerc.pdf}
			\caption{Charges percentage.}
			\label{fig:7_7a_recharge_Needed}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/AvgSOC_comparison_N}
			\caption{Average state of charge.}
			\label{fig:7_7a_asoc_Needed}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Needed_ReroutePerc.pdf}
			\caption{Rerouted trips percentage.}
			\label{fig:7_7a_reroute_Needed}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Needed_AvgWalkedDistance.pdf}
			\caption{Walked distance when rerouted.}
			\label{fig:7_7a_awd_Needed}
		\end{subfigure}         
		\caption{\textit{Genetic} and \textit{local search} optimization results for metrics of interests (\textit{Needed} return policy adopted).}
		\label{fig:7_7a_opt_needed}
	\end{center}
\end{figure*}

Figure~\ref{fig:7_7a_opt_needed} reports the other user discomfort metrics. The two optimisation algorithms reduce the charge events (Fig~\ref{fig:7_7a_recharge_Needed}).
%The \textit{Genetic} algorithm shows the best performance with a probability the lowest probability of 6\% when 13\% of the zones are equipped with charging stations. 
Since the car are recharged less frequently with respect to the \textit{Hybrid} policy it is interesting to evaluate the impact on the \textit{Average Stage of Charge}. figure~\ref{fig:7_7a_asoc_Needed} reports this metric for the different placement algorithms.  As we expected, the \textit{Average Stage of Charge} is lower with respect to the \textit{Hybrid policy}. Interestingly, for both optimized solutions this value saturate almost immediately at 55\% in the feasible region.
% Moreover, despite in figure~\ref{fig:7_7a_recharge_Needed} we saw that in the optimized solutions we recharge less frequent the car, 
However,  the average state of charge (figure~\ref{fig:7_7a_asoc_Needed}) is always higher in the optimised solutions than in the \textit{Num parking} configuration. This further demonstrates that a smarter placement allows the car to get more energy for each charge.
%with an opposite trend with respect to Fig~\ref{fig:7_7a_recharge_Hybrid},
%Finally, we evaluate the users discomfort metrics in terms of rerouting probability and the average distance an user has to walk when rerouted.
Figure~\ref{fig:7_7a_reroute_Needed} reports the re-routing percentage for the different algorithms. As expected, the values are larger than with the \textit{Hybrid} policy, since no opportunistic charge is performed. 
This different is particularly evident when a few charging stations are present e.g., with 5\% of charging stations the rerouting are 14\% for the \textit{Num parking} heuristic, 14\% for the \textit{Local search}, and 6\% for the \textit{Genetic} algorithm. 
However, the genetic algorithm is able to quickly reach a small value of re-routings, hence better exploiting every charge possibility. 
%without showing further improvements while increasing the percentage of charging stations applied. 
Finally, I analyse the distance a user has to walk when rerouted (Figure~\ref{fig:7_7a_awd_Needed}). Here, with respect to the \textit{Hybrid} policy case, the \textit{local search} shows larger deviations from the \textit{Num parking}. The genetic algorithm reaches values of about 800\,m, even below the ones reached with the \textit{Hybrid} policy.

>>>>

 \begin{figure}[h]
     \centering     %%% not \center
    \subfloat[][Percentage of infeasible trips. Y-Axis is logarithmic.]
    {
        \includegraphics[width=0.45\textwidth]{figures/Needed_Deaths.pdf}
        \label{fig:7_7b_optimized_deaths_Needed}
    }     
     \subfloat[][Walked distance, averaged over all trips.]
     {
         \includegraphics[width=0.45\textwidth]{figures/Needed_TravelWithPenlaty.pdf}
         \label{fig:7_7b_wwd_Needed}
     }
     \caption{Objective metrics to minimise in the optimisation - with \textit{Needed} return policy}
    \label{fig:7_7b_optimized_metrics_needed}
 \end{figure}

Here I briefly report the results for the optimisation experiments of the \textit{Needed} policy.
I followed the same procedure explained in Sec.~\ref{sec:7_7a_opt} for the \textit{Hybrid} return policy. As in that case, the genetic algorithm is able to largely optimise the solution, as reported in figure~\ref{fig:7_7b_optimized_metrics_needed}, with \textit{local searches} stuck in local minima. 
In particular, for the walked distance (figure~\ref{fig:7_7b_wwd_Needed}), the genetic algorithm is able lower it from 136\,m to 45\,m at 13\% of zones. Still, it doesn't reach the performance of the \textit{Hybrid} policy, i.e., 30\,m at 13\% of zones.

\begin{figure*}[t!]
	\begin{center}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Needed_AmountRechargePerc.pdf}
			\caption{Charges percentage.}
			\label{fig:7_7b_recharge_Needed}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/AvgSOC_comparison_N}
			\caption{Average state of charge.}
			\label{fig:7_7b_asoc_Needed}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Needed_ReroutePerc.pdf}
			\caption{Rerouted trips percentage.}
			\label{fig:7_7b_reroute_Needed}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\includegraphics[width=\columnwidth]{figures/Needed_AvgWalkedDistance.pdf}
			\caption{Walked distance when rerouted.}
			\label{fig:7_7b_awd_Needed}
		\end{subfigure}         
		\caption{\textit{Genetic} and \textit{local search} optimization results for metrics of interests (\textit{Needed} return policy adopted).}
		\label{fig:7_7b_opt_needed}
	\end{center}
\end{figure*}

Figure~\ref{fig:7_7b_opt_needed} reports the other user discomfort metrics. The two optimisation algorithms reduce the charge events (Fig~\ref{fig:7_7b_recharge_Needed}).
%The \textit{Genetic} algorithm shows the best performance with a probability the lowest probability of 6\% when 13\% of the zones are equipped with charging stations. 
Since the car are recharged less frequently with respect to the \textit{Hybrid} policy it is interesting to evaluate the impact on the \textit{Average Stage of Charge}. figure~\ref{fig:7_7b_asoc_Needed} reports this metric for the different placement algorithms.  As we expected, the \textit{Average Stage of Charge} is lower with respect to the \textit{Hybrid policy}. Interestingly, for both optimized solutions this value saturate almost immediately at 55\% in the feasible region.
% Moreover, despite in figure~\ref{fig:7_7b_recharge_Needed} we saw that in the optimized solutions we recharge less frequent the car, 
However,  the average state of charge (figure~\ref{fig:7_7b_asoc_Needed}) is always higher in the optimised solutions than in the \textit{Num parking} configuration. This further demonstrates that a smarter placement allows the car to get more energy for each charge.
%with an opposite trend with respect to Fig~\ref{fig:7_7b_recharge_Hybrid},
%Finally, we evaluate the users discomfort metrics in terms of rerouting probability and the average distance an user has to walk when rerouted.
Figure~\ref{fig:7_7b_reroute_Needed} reports the re-routing percentage for the different algorithms. As expected, the values are larger than with the \textit{Hybrid} policy, since no opportunistic charge is performed. 
This different is particularly evident when a few charging stations are present e.g., with 5\% of charging stations the rerouting are 14\% for the \textit{Num parking} heuristic, 14\% for the \textit{Local search}, and 6\% for the \textit{Genetic} algorithm. 
However, the genetic algorithm is able to quickly reach a small value of re-routings, hence better exploiting every charge possibility. 
%without showing further improvements while increasing the percentage of charging stations applied. 
Finally, I analyse the distance a user has to walk when rerouted (Figure~\ref{fig:7_7b_awd_Needed}). Here, with respect to the \textit{Hybrid} policy case, the \textit{local search} shows larger deviations from the \textit{Num parking}. The genetic algorithm reaches values of about 800\,m, even below the ones reached with the \textit{Hybrid} policy.

### 2020-11-12 Chapter7/7_8_conclusion.tex
chapt

\subsection{\textbf{Scalability}}

@@

Directly linked to the scalability problem, another important aspect is the possibility to add new charging stations when car sharing demand changes. 
By analysing the data collected by UMAP, this methodology can be used to consider the greedy placement of new charging stations on the top of those already present.

\subsection{\textbf{Economical Aspects}}

>>>>

\subsection{Scalability}

@@

Directly linked to the scalability problem, another important aspect is the possibility to add new charging stations when car sharing demand changes. 
By analysing the data collected, this methodology can be used to consider the greedy placement of new charging stations on the top of those already present.

\subsection{Economical Aspects}

### 2020-11-12 Chapter8/8_1_introduction.tex
chapt

Car sharing is an evolution of the classic car rental model. Here, users can rent cars on-demand for a short period, e.g., a twenty-minute trip across town. In particular, Free Floating Car Sharing (FFCS) services allow customers to rent and return the cars everywhere inside a operative area in a city. Customers book, unlock and return the car by using an application on their smartphones. One such service is Car2go~\footnote{\url{https://www.car2go.com/}}, which currently operates in several cities around the world. In the FFCS implementation, of which Car2go is an example, the provider bills the user only for the time spent driving, with simple minute-based fares which factors all costs. Some studies demonstrate that a massive adoption of car sharing service can improve mobility as well as reduce costs and pollution~\citep{EvaluatingCarsharingBenefits,firnkorn2011,FM12}.

>>>>

Car sharing is an evolution of the classic car rental model. Here, users can rent cars on-demand for a short period, e.g., a twenty-minute trip across town. In particular, Free Floating Car Sharing (FFCS) services allow customers to rent and return the cars everywhere inside a operative area in a city. Customers book, unlock and return the car by using an application on their smartphones. One such service is Car2go~\footnote{\url{https://www.car2go.com/}}, which currently operates in several cities around the world. In the FFCS implementation, of which Car2go is an example, the provider bills the user only for the time spent driving, with simple minute-based fares which factors all costs. Some studies demonstrate that a massive adoption of car sharing service can improve mobility as well as reduce costs and pollution~\citep{EvaluatingCarsharingBenefits,Firnkorn2011,FM12}.

### 2020-11-12 Chapter8/8_2_related_works.tex
chapt

% Related works on traffic demand modelling  and prediction
With the easiness of collecting data and the ability to build and train off-the-shelf machine learning solutions, researchers have started applying data driven approaches in the context of transportation.
Previous work~\citep{2_okutani1984dynamic} addressed traffic modelling and prediction with real traffic data, and proposes strategies to improve congestion prediction using Kalman filters, showing how traffic is stationary in time.  
Other studies~\citep{3_clark2003traffic} proposed new approaches, based on a multivariate extension of non-parametric regression, to predict traffic patterns, with the goal of counteracting traffic congestion. 
While similar in spirit, our work focuses on FFCS services explicitly, and uses a much richer dataset as well as more advanced machine learning algorithms.

@@

Other studies~\citep{schmoller2014analyzing,Schmoller2015} conducted a detailed characterization of a car sharing system in Munich and Berlin. Similarly to our work, they identified features correlated with the demand for shared cars in the target cities. However, our work differs from their in the sense that we here analyze a much larger set of features, including demographics and economic data, and consider multiple prediction models. We focus on demand prediction, facing both time and space dimensions, and provide a thorough comparison and guidelines for future directions.

In our previous work~\citep{VancouverCS}, we analyzed in depth the usage of different car sharing systems in Vancouver. Based on this data, we developed a model of FFCS usage and built a simulator to design new systems based on electric vehicles~\citep{coccacar}. In particular, we tackled the charging station placement problem, showing that the optimal placement requires few stations to satisfy charging requests in different cities~\citep{coccaopt}.

@@

Furthermore, while previous work~\citep{wang2017deepsd} focused on the temporal prediction of car sharing demand in a very short-term basis (demand prediction in the next few minutes), in this work we focus on the problem at different time scales. We also compare several prediction strategies and analyze how the temporal prediction problem relates to the spatial prediction one.
Moreover, we are the first to use a very heterogeneous dataset including dozens of features to tackle the prediction problems. This allows us to provide insights on which of those features are the most important ones to solve our prediction problems as well as to have a broader perspective on the challenges involved in car sharing prediction.
%}

>>>>

% Related works on traffic demand modelling  and prediction
With the easiness of collecting data and the ability to build and train off-the-shelf machine learning solutions, researchers have started applying data driven approaches in the context of transportation.
Previous work~\citep{2_okutani1984dynamic} addressed traffic modelling and prediction with real traffic data, and proposes strategies to improve congestion prediction using Kalman filters, showing how traffic is stationary in time.  
Other studies~\citep{3_clark2003traffic} proposed new approaches, based on a multivariate extension of non-parametric regression, to predict traffic patterns, with the goal of counteracting traffic congestion. 
While similar in spirit, this work focuses on FFCS services explicitly, and uses a much richer dataset as well as more advanced machine learning algorithms.

@@

Other studies~\citep{schmoller2014analyzing,Schmoller2015} conducted a detailed characterization of a car sharing system in Munich and Berlin. Similarly to this work, they identified features correlated with the demand for shared cars in the target cities. However, this work differs from their in the sense that we here analyze a much larger set of features, including demographics and economic data, and consider multiple prediction models. We focus on demand prediction, facing both time and space dimensions, and provide a thorough comparison and guidelines for future directions.

In this previous work~\citep{VancouverCS}, we analyzed in depth the usage of different car sharing systems in Vancouver. Based on this data, we developed a model of FFCS usage and built a simulator to design new systems based on electric vehicles~\citep{coccacar}. In particular, we tackled the charging station placement problem, showing that the optimal placement requires few stations to satisfy charging requests in different cities~\citep{coccaopt}.

@@

Furthermore, while previous work~\citep{wang2017deepsd} focused on the temporal prediction of car sharing demand in a very short-term basis (demand prediction in the next few minutes), in this work we focus on the problem at different time scales. We also compare several prediction strategies and analyze how the temporal prediction problem relates to the spatial prediction one.
Moreover, we are the first to use a very heterogeneous dataset including dozens of features to tackle the prediction problems. This allows us to provide insights on which of those features are the most important ones to solve the prediction problems as well as to have a broader perspective on the challenges involved in car sharing prediction.
%}

### 2020-11-12 Chapter8/8_4_temporal_analyses.tex
chapt

\begin{table*}
\centering
\begin{tabular}{@{}rrrrcrrrr@{}}
\toprule
& \multicolumn{3}{c}{Expanding Window (starting: 28 days)} & \phantom{abc} & \multicolumn{3}{c}{Sliding Window (28 days)} \\
\cmidrule{2-4} \cmidrule{6-8}
 {Prediction Model} & \textbf{MAPE [\%]} & \textbf{$\sigma$(APE) [\%]} & \textbf{RMSE} && \textbf{MAPE [\%]} & \textbf{$\sigma$(APE) [\%]} & \textbf{RMSE} \\ 
 %& \textbf{Average} & \textbf{Std.  Dev.} & \textbf{Average} && \textbf{Average} & \textbf{Std. Dev.} & \textbf{Average} \\ 
 \midrule
  \textbf{Baseline}                     & 20.12 & 16.64 & 195.53 && 20.12 & 16.64 & 195.53 \\
  \textbf{ARIMA}                        & 36.01 & 35.87 & 306.80 && 36.52 & 36.60 & 305.50 \\
  \textbf{SARIMA}                       & 17.60 & 20.01 & 160.42 && 18.02 & 21.75 & 163.94 \\
  \textbf{Linear Regression}            & 18.28 & 20.38 & 179.11 && 18.11 & 20.55 & 178.61 \\
  \textbf{Support Vector Regression}    & 12.22 & 15.62 & 128.72 && 12.87 & 18.52 & 136.14 \\  
  \textbf{Random Forests Regression}    & 9.71  & 8.34  & 104.99 && 10.08 & 12.23 & 109.47 \\
  \textbf{Neural Networks}              & 10.52 & 12.93 & 128.84 && 10.52 & 12.74 & 123.55 \\
\bottomrule
\end{tabular}
\caption{Short-term temporal prediction - Mean Absolute Percentage Error (MAPE), Absolute Percentage Error (APE), and Root Mean Squared Error (RMSE), for each prediction model in the validation set.}
\label{tab:8_4_results-dynamic-models}
\end{table*}

>>>>

%\begin{table*}
%	\centering
%	\begin{tabular}{@{}rrrrcrrrr@{}}
%	\toprule
%	& \multicolumn{3}{c}{Expanding Window (starting: 28 days)} & \phantom{abc} & \multicolumn{3}{c}{Sliding Window (28 days)} \\
%	\cmidrule{2-4} \cmidrule{6-8}
%	 {Prediction Model} & \textbf{MAPE [\%]} & \textbf{$\sigma$(APE) [\%]} & \textbf{RMSE} && \textbf{MAPE [\%]} & \textbf{$\sigma$(APE) [\%]} & \textbf{RMSE} \\ 
%	 %& \textbf{Average} & \textbf{Std.  Dev.} & \textbf{Average} && \textbf{Average} & \textbf{Std. Dev.} & \textbf{Average} \\ 
%	 \midrule
%	  \textbf{Baseline}                     & 20.12 & 16.64 & 195.53 && 20.12 & 16.64 & 195.53 \\
%	  \textbf{ARIMA}                        & 36.01 & 35.87 & 306.80 && 36.52 & 36.60 & 305.50 \\
%	  \textbf{SARIMA}                       & 17.60 & 20.01 & 160.42 && 18.02 & 21.75 & 163.94 \\
%	  \textbf{Linear Regression}            & 18.28 & 20.38 & 179.11 && 18.11 & 20.55 & 178.61 \\
%	  \textbf{Support Vector Regression}    & 12.22 & 15.62 & 128.72 && 12.87 & 18.52 & 136.14 \\  
%	  \textbf{Random Forests Regression}    & 9.71  & 8.34  & 104.99 && 10.08 & 12.23 & 109.47 \\
%	  \textbf{Neural Networks}              & 10.52 & 12.93 & 128.84 && 10.52 & 12.74 & 123.55 \\
%	\bottomrule
%	\end{tabular}
%	\caption{Short-term temporal prediction - Mean Absolute Percentage Error (MAPE), Absolute Percentage Error (APE), and Root Mean Squared Error (RMSE), for each prediction model in the validation set.}
%\label{tab:8_4_results-dynamic-models}
%\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{@{}rrrrcrrrr@{}}
	\toprule\toprule
	& \multicolumn{3}{c}{Expanding Window (starting: 28 days)} & \phantom{abc} \\
	\cmidrule{2-4} \cmidrule{6-8}
	 {Prediction Model} & \textbf{MAPE [\%]} & \textbf{$\sigma$(APE) [\%]} & \textbf{RMSE}\\ 
	 %& \textbf{Average} & \textbf{Std.  Dev.} & \textbf{Average} && \textbf{Average} & \textbf{Std. Dev.} & \textbf{Average} \\ 
	 \midrule[0.5pt]
	  \textbf{Baseline}                     & 20.12 & 16.64 & 195.53\\
	  \textbf{ARIMA}                        & 36.01 & 35.87 & 306.80\\
	  \textbf{SARIMA}                       & 17.60 & 20.01 & 160.42\\
	  \textbf{Linear Regression}            & 18.28 & 20.38 & 179.11\\
	  \textbf{Support Vector Regression}    & 12.22 & 15.62 & 128.72\\  
	  \textbf{Random Forests Regression}    & 9.71  & 8.34  & 104.99\\
	  \textbf{Neural Networks}              & 10.52 & 12.93 & 128.84\\

	  \toprule

	  	& \multicolumn{3}{c}{Sliding Window (starting: 28 days)} & \phantom{abc} \\
	  \cmidrule{2-4} \cmidrule{6-8}
	  {Prediction Model} & \textbf{MAPE [\%]} & \textbf{$\sigma$(APE) [\%]} & \textbf{RMSE}\\ 
	  %& \textbf{Average} & \textbf{Std.  Dev.} & \textbf{Average} && \textbf{Average} & \textbf{Std. Dev.} & \textbf{Average} \\ 
	  \midrule
	  \textbf{Baseline}                     & 20.12 & 16.64 & 195.53 \\
	  \textbf{ARIMA}                        & 36.52 & 36.60 & 305.50 \\
	  \textbf{SARIMA}                       & 18.02 & 21.75 & 163.94 \\
	  \textbf{Linear Regression}            & 18.11 & 20.55 & 178.61 \\
	  \textbf{Support Vector Regression}    & 12.87 & 18.52 & 136.14 \\  
	  \textbf{Random Forests Regression}    & 10.08 & 12.23 & 109.47 \\
	  \textbf{Neural Networks}              & 10.52 & 12.74 & 123.55 \\

	\bottomrule\bottomrule
	\end{tabular}
	\caption{Short-term temporal prediction - Mean Absolute Percentage Error (MAPE), Absolute Percentage Error (APE), and Root Mean Squared Error (RMSE), for each prediction model in the validation set.}
\label{tab:8_4_results-dynamic-models}
\end{table*}
